---
title: "Pose-Estimation Pipeline — Analysis Summary"
author: "Generated from analysis/*.R"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    df_print: paged
    fig_width: 10
    fig_height: 6
params:
  output_dir: "../output"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo       = FALSE,
  warning    = FALSE,
  message    = FALSE,
  fig.align  = "center",
  out.width  = "100%"
)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(readr)
  library(stringr)
  library(purrr)
  library(ggplot2)
  library(tibble)
  library(jsonlite)
  library(knitr)
})

out_dir <- normalizePath(params$output_dir, mustWork = FALSE)
has_out <- dir.exists(out_dir)

show_png <- function(path, caption = NULL) {
  if (file.exists(path)) {
    knitr::include_graphics(path)
  } else {
    cat(sprintf("_Not available: `%s`_\n\n", basename(path)))
  }
}

read_csv_safe <- function(path) {
  if (file.exists(path)) {
    suppressMessages(read_csv(path, show_col_types = FALSE))
  } else {
    NULL
  }
}

list_files_safe <- function(dir, pattern) {
  if (dir.exists(dir)) {
    list.files(dir, pattern = pattern, full.names = TRUE)
  } else {
    character()
  }
}
```

# Overview

This report consolidates the outputs of the R analysis scripts in
`analysis/` that consume the CSVs produced by the pose-estimation
pipeline (`pose_estimation.main` / `pose_estimation.run`). It is
organised in the order the scripts are typically run:

1. **Pipeline quality** — `summary.R`, `timeseries.R`, `keypoint_detail.R`
2. **Landmark feature exploration** — `features.R`
3. **Run comparison** — `compare.R`
4. **Clinical kinematic features** — `clinical_features.R`
5. **Clinical exploration** — `explore_clinical.R`
6. **Temporal clinical analysis** — `temporal_clinical.R`
7. **Between-video comparison** — `compare_clinical.R`
8. **Dimensionality reduction** — `clinical_dimreduce.R`
9. **External-score correlation** — `clinical_correlation.R`
10. **Longitudinal recovery** — `longitudinal.R`
11. **Metadata templates & validation** — `make_templates.R`, `validate_metadata.R`

The report reads whatever outputs already exist under
`r out_dir` and silently skips sections whose inputs have not been
produced yet.

```{r inventory}
count_landmark_csvs <- function(dir) {
  if (!dir.exists(dir)) return(NA_integer_)
  files <- list.files(dir, pattern = "\\.csv$", full.names = TRUE)
  excl <- "(_metrics|_diag|_smooth|_clinical|_kp_detail|_summary|_feature_rank)\\.csv$"
  length(files[!str_detect(basename(files), excl)])
}

count_pattern <- function(dir, pattern, extra_exclude = NULL) {
  if (!dir.exists(dir)) return(NA_integer_)
  files <- list.files(dir, pattern = pattern, full.names = TRUE)
  if (!is.null(extra_exclude)) {
    files <- files[!str_detect(basename(files), extra_exclude)]
  }
  length(files)
}

inventory <- tibble(
  category = c(
    "Landmark CSVs (pipeline export)",
    "Metrics CSVs (quality metrics)",
    "Per-keypoint detail CSVs",
    "Per-run summary JSONs",
    "Per-frame clinical CSVs",
    "Per-window clinical CSVs"
  ),
  n_files = c(
    count_landmark_csvs(out_dir),
    count_pattern(out_dir, "_metrics\\.csv$"),
    count_pattern(out_dir, "_kp_detail\\.csv$"),
    count_pattern(out_dir, "_summary\\.json$"),
    count_pattern(out_dir, "_clinical\\.csv$", extra_exclude = "_windows\\.csv$"),
    count_pattern(out_dir, "_clinical_windows\\.csv$")
  )
)

knitr::kable(
  inventory |> select(category, n_files),
  col.names = c("Output category", "# files"),
  caption = "Output inventory"
)
```

---

# Pipeline Quality

## Per-run baseline report — `summary.R`

`summary.R` ingests each `*_metrics.csv` and produces a JSON summary
(`*_summary.json`) plus a console report covering:

- Detection rates (body, synthetic hands, re-crop hands)
- Confidence (body visibility median/IQR, hand flag medians)
- Jitter (body/hand median and p95, px)
- Smoothing delta (raw − smoothed, px)
- Carry-forward rate and mean streak length
- Bone-length correction and joint-angle clamp rate
- Hand-to-arm matching distances
- Inference time (median and p95, ms)

```{r summary-json-table}
json_files <- list_files_safe(out_dir, "_summary\\.json$")

if (length(json_files) == 0) {
  cat("_No `*_summary.json` files found. Run:_ `Rscript analysis/summary.R output/`\n")
} else {
  flatten_json <- function(j, prefix = "") {
    out <- list()
    for (name in names(j)) {
      key <- if (prefix == "") name else paste0(prefix, ".", name)
      val <- j[[name]]
      if (is.list(val) && !is.null(names(val))) {
        out <- c(out, flatten_json(val, key))
      } else {
        out[[key]] <- val
      }
    }
    out
  }

  rows <- map_dfr(json_files, \(f) {
    j <- jsonlite::fromJSON(f)
    flat <- flatten_json(j)
    video <- str_remove(basename(f), "_summary\\.json$")
    tibble(video = video, !!!flat)
  })

  knitr::kable(
    rows |>
      select(video, n_frames,
             any_of(c(
               "detection.body_detection_rate",
               "confidence.body_vis_median",
               "jitter_px.body_median",
               "jitter_px.body_p95",
               "carry_forward.body_carry_pct",
               "inference_ms.median",
               "inference_ms.p95"
             ))),
    caption = "Per-video baseline quality (from `*_summary.json`)"
  )
}
```

## Temporal diagnostics — `timeseries.R`

For every `*_metrics.csv`, `timeseries.R` generates five diagnostic
plots:

- `_detection_timeline.png` — frame-by-frame body/hand status
  (detected / carried / missing)
- `_jitter.png` — jitter over time with spike markers (> 3 × median)
- `_confidence.png` — body visibility + hand flag traces
- `_smoothing.png` — raw-vs-smoothed delta
- `_constraints.png` — bone-length and joint-angle correction history

```{r timeseries-examples, results = "asis"}
timeseries_pngs <- list_files_safe(out_dir, "_detection_timeline\\.png$")

if (length(timeseries_pngs) == 0) {
  cat("_No timeseries plots found. Run:_ `Rscript analysis/timeseries.R output/`\n")
} else {
  # Show the first example per plot type
  example <- str_remove(basename(timeseries_pngs[1]), "_detection_timeline\\.png$")
  cat(sprintf("Showing diagnostic plots for **%s** (one of %d videos):\n\n",
              example, length(timeseries_pngs)))
  for (suffix in c("_detection_timeline", "_jitter", "_confidence",
                   "_smoothing", "_constraints")) {
    p <- file.path(out_dir, paste0(example, suffix, ".png"))
    if (file.exists(p)) {
      cat(sprintf("### `%s`\n\n", basename(p)))
      cat(sprintf("![](%s)\n\n", p))
    }
  }
}
```

## Per-keypoint analysis — `keypoint_detail.R`

Runs when `*_kp_detail.csv` files are exported
(`--metrics-detail` flag). Produces:

- `_jitter_heatmap.png` — keypoint × time jitter heatmap
- `_worst_keypoints.png` — top-20 keypoints by median jitter
- `_trajectory.png` — smoothed x/y trajectories for shoulders + hand wrists

```{r keypoint-detail, results = "asis"}
kp_pngs <- list_files_safe(out_dir, "_jitter_heatmap\\.png$")

if (length(kp_pngs) == 0) {
  cat("_No per-keypoint plots found. Export with `--metrics-detail`, then run:_ ",
      "`Rscript analysis/keypoint_detail.R output/`\n")
} else {
  example <- str_remove(basename(kp_pngs[1]), "_jitter_heatmap\\.png$")
  cat(sprintf("Showing per-keypoint plots for **%s** (one of %d videos):\n\n",
              example, length(kp_pngs)))
  for (suffix in c("_worst_keypoints", "_jitter_heatmap", "_trajectory")) {
    p <- file.path(out_dir, paste0(example, suffix, ".png"))
    if (file.exists(p)) {
      cat(sprintf("### `%s`\n\n![](%s)\n\n", basename(p), p))
    }
  }
}
```

---

# Landmark Feature Exploration — `features.R`

`features.R` loads landmark CSVs (`hands-arms` or `body` tracking
mode), drops rows that are entirely NA, median-imputes sporadic NAs,
and produces:

- `_variance.png` — per-feature variance (log scale), coloured by part
- `_correlation.png` — clustered Pearson correlation heatmap
- `_pca_scree.png` — cumulative variance explained
- `_pca_biplot.png` — PC1 vs PC2 with top-10 loading arrows,
  points coloured by timestamp
- `_umap.png` / `_umap_video.png` — 2D UMAP embeddings
- `_feature_rank.csv` — variance / max-abs-correlation / top-PC loading table

```{r features-rank-table}
rank_csvs <- list_files_safe(out_dir, "_feature_rank\\.csv$")

if (length(rank_csvs) == 0) {
  cat("_No `*_feature_rank.csv` found. Run:_ `Rscript analysis/features.R output/`\n")
} else {
  rank <- read_csv_safe(rank_csvs[1])
  if (!is.null(rank)) {
    cat(sprintf("Top 10 features by variance (from `%s`):\n\n",
                basename(rank_csvs[1])))
    knitr::kable(
      rank |> arrange(desc(variance)) |> head(10) |>
        select(any_of(c("column", "part", "keypoint", "coord",
                        "variance", "max_abs_corr", "top_pc", "pc_loading"))),
      digits = 3
    )
  }
}
```

```{r features-plots, results = "asis"}
var_pngs <- list_files_safe(out_dir, "_variance\\.png$")
if (length(var_pngs) > 0) {
  example <- str_remove(basename(var_pngs[1]), "_variance\\.png$")
  cat(sprintf("Plots for **%s**:\n\n", example))
  for (suffix in c("_variance", "_correlation", "_pca_scree",
                   "_pca_biplot", "_umap")) {
    p <- file.path(out_dir, paste0(example, suffix, ".png"))
    if (file.exists(p)) {
      cat(sprintf("![](%s)\n\n", p))
    }
  }
}
```

---

# Run Comparison — `compare.R`

`compare.R` diffs two pipeline runs (`*_summary.json` or
`*_metrics.csv`), writing `comparison.csv` plus — when metrics CSVs
are provided — `comparison_jitter.png` and
`comparison_detection.png` (overlaid distributions and bar chart).

```{r compare-table}
comp_csv <- file.path(out_dir, "comparison.csv")
comp <- read_csv_safe(comp_csv)
if (is.null(comp)) {
  cat("_No `comparison.csv` found. Run:_ ",
      "`Rscript analysis/compare.R run_a.json run_b.json`\n")
} else {
  knitr::kable(comp |> arrange(desc(abs(pct_change))) |> head(15),
               caption = "Top 15 metrics by |percentage change|",
               digits = 3)
}
```

```{r compare-plots, results = "asis"}
for (f in c("comparison_jitter.png", "comparison_detection.png")) {
  p <- file.path(out_dir, f)
  if (file.exists(p)) cat(sprintf("![](%s)\n\n", p))
}
```

---

# Clinical Kinematic Features — `clinical_features.R`

Per-frame features (joint angles, reach, grasp aperture, wrist /
fingertip displacement) and per-window (1 s, 50 % overlap)
smoothness features (spectral arc length, wrist velocity
mean / peak) are derived from the landmark CSVs. Two files are
written per input:

- `<stem>_clinical.csv` — per-frame features
- `<stem>_clinical_windows.csv` — per-window features

```{r clinical-features-inventory}
frame_files <- list_files_safe(out_dir, "_clinical\\.csv$")
frame_files <- frame_files[!str_detect(basename(frame_files), "_windows\\.csv$")]
win_files   <- list_files_safe(out_dir, "_clinical_windows\\.csv$")

if (length(frame_files) == 0) {
  cat("_No `*_clinical.csv` found. Run:_ ",
      "`Rscript analysis/clinical_features.R output/`\n")
} else {
  inv <- tibble(
    video = str_remove(basename(frame_files), "_clinical\\.csv$"),
    frame_rows = map_int(frame_files, \(f) nrow(read_csv_safe(f)))
  )
  if (length(win_files) > 0) {
    inv <- inv |>
      left_join(
        tibble(
          video = str_remove(basename(win_files), "_clinical_windows\\.csv$"),
          window_rows = map_int(win_files, \(f) nrow(read_csv_safe(f)))
        ),
        by = "video"
      )
  } else {
    inv$window_rows <- NA_integer_
  }
  knitr::kable(inv, caption = "Clinical feature files (per video)")
}
```

---

# Clinical Exploration — `explore_clinical.R`

Loads all per-frame and per-window clinical CSVs and writes four
diagnostic plots plus a console summary of row counts, per-column
NA rates, and data-quality warnings (features > 50 % NA,
zero-variance columns, empty videos).

```{r explore-plots, results = "asis"}
for (f in c("all_clinical_distributions.png",
            "all_clinical_na_heatmap.png",
            "all_clinical_boxplots.png",
            "all_clinical_window_distributions.png")) {
  p <- file.path(out_dir, f)
  if (file.exists(p)) {
    cat(sprintf("### `%s`\n\n![](%s)\n\n", f, p))
  }
}
```

---

# Temporal Clinical Analysis — `temporal_clinical.R`

For every video with ≥ 10 per-frame rows, `temporal_clinical.R`
produces a multi-panel time-series plot with left and right sides
overlaid, plus a window-level SAL / velocity panel where
available. A cross-video overview plot compares normalised reach
across all qualifying videos.

```{r temporal-plots, results = "asis"}
ts_pngs <- list_files_safe(out_dir, "_clinical_timeseries\\.png$")
overview <- file.path(out_dir, "all_clinical_timeseries_overview.png")

if (file.exists(overview)) {
  cat("### Cross-video overview\n\n")
  cat(sprintf("![](%s)\n\n", overview))
}

if (length(ts_pngs) > 0) {
  cat(sprintf("### Example per-video time series (%d total)\n\n",
              length(ts_pngs)))
  cat(sprintf("![](%s)\n\n", ts_pngs[1]))
}

if (length(ts_pngs) == 0 && !file.exists(overview)) {
  cat("_No temporal plots found. Run:_ ",
      "`Rscript analysis/temporal_clinical.R output/`\n")
}
```

---

# Between-Video Comparison — `compare_clinical.R`

Aggregates per-frame clinical features per video (mean / median /
SD / min / max) and window features (mean / SD), producing a wide
summary table (`all_clinical_video_summary.csv`) and two plots:

- `all_clinical_radar.png` — parallel-coordinate plot of
  z-scored feature means
- `all_clinical_heatmap.png` — clustered z-score heatmap

Outlier videos (|z| > 2 on any feature mean) are printed to the
console.

```{r compare-clinical-summary}
video_summary <- read_csv_safe(file.path(out_dir, "all_clinical_video_summary.csv"))

if (is.null(video_summary)) {
  cat("_No `all_clinical_video_summary.csv` found. Run:_ ",
      "`Rscript analysis/compare_clinical.R output/`\n")
} else {
  cat(sprintf("**%d videos × %d aggregated feature columns.**\n\n",
              nrow(video_summary), ncol(video_summary) - 1))

  mean_cols <- names(video_summary)[str_ends(names(video_summary), "__mean")]
  if (length(mean_cols) > 0) {
    z_df <- video_summary |>
      select(video, all_of(mean_cols)) |>
      mutate(across(all_of(mean_cols), \(x) {
        s <- sd(x, na.rm = TRUE)
        if (is.na(s) || s == 0) return(rep(NA_real_, length(x)))
        (x - mean(x, na.rm = TRUE)) / s
      }))

    outliers <- z_df |>
      pivot_longer(-video, names_to = "feature", values_to = "z") |>
      filter(!is.na(z), abs(z) > 2) |>
      mutate(feature = str_remove(feature, "__mean$")) |>
      arrange(desc(abs(z)))

    if (nrow(outliers) > 0) {
      knitr::kable(outliers |> head(20),
                   digits = 2,
                   caption = "Outlier videos (|z| > 2 on any feature mean, top 20)")
    } else {
      cat("_No videos exceed |z| > 2 on any feature._\n\n")
    }
  }
}
```

```{r compare-clinical-plots, results = "asis"}
for (f in c("all_clinical_radar.png", "all_clinical_heatmap.png")) {
  p <- file.path(out_dir, f)
  if (file.exists(p)) {
    cat(sprintf("### `%s`\n\n![](%s)\n\n", f, p))
  }
}
```

---

# Dimensionality Reduction — `clinical_dimreduce.R`

PCA and UMAP on the per-video aggregated clinical features.
Writes:

- `all_clinical_pca_scree.png` — variance explained per PC
- `all_clinical_pca_biplot.png` — PC1/PC2 scatter with top loading arrows
- `all_clinical_umap.png` — 2D UMAP
- `all_clinical_pca_loadings.csv` — feature loadings on first 5 PCs

```{r dimreduce-loadings}
loadings <- read_csv_safe(file.path(out_dir, "all_clinical_pca_loadings.csv"))

if (is.null(loadings)) {
  cat("_No PCA loadings found. Run:_ ",
      "`Rscript analysis/clinical_dimreduce.R output/`\n")
} else {
  knitr::kable(
    loadings |> head(10) |>
      select(any_of(c("feature_short", "PC1", "PC2", "PC3", "abs_PC1"))),
    digits = 3,
    caption = "Top 10 features loading on PC1"
  )
}
```

```{r dimreduce-plots, results = "asis"}
for (f in c("all_clinical_pca_scree.png",
            "all_clinical_pca_biplot.png",
            "all_clinical_umap.png")) {
  p <- file.path(out_dir, f)
  if (file.exists(p)) {
    cat(sprintf("### `%s`\n\n![](%s)\n\n", f, p))
  }
}
```

---

# External-Score Correlation — `clinical_correlation.R`

Joins aggregated clinical features (mean / median / SD / min / max
per video) with an external `clinical_scores.csv` on `video` and
computes pairwise Pearson and Spearman correlations with
Benjamini–Hochberg FDR correction. Outputs:

- `clinical_correlation_table.csv` — tidy correlation results
- `clinical_correlation_matrix.png` — Spearman heatmap with
  significance stars
- `clinical_scatter_top.png` — scatter plots for top 6 pairs

```{r correlation-table}
cor_tbl <- read_csv_safe(file.path(out_dir, "clinical_correlation_table.csv"))

if (is.null(cor_tbl)) {
  cat("_No `clinical_correlation_table.csv` found. Run:_ ",
      "`Rscript analysis/clinical_correlation.R output/ clinical_scores.csv`\n")
} else {
  top <- cor_tbl |>
    filter(!is.na(spearman_rho)) |>
    arrange(desc(abs(spearman_rho))) |>
    head(15)
  knitr::kable(top, digits = 3,
               caption = "Top 15 feature–score pairs by |Spearman ρ|")
}
```

```{r correlation-plots, results = "asis"}
for (f in c("clinical_correlation_matrix.png", "clinical_scatter_top.png")) {
  p <- file.path(out_dir, f)
  if (file.exists(p)) {
    cat(sprintf("### `%s`\n\n![](%s)\n\n", f, p))
  }
}
```

---

# Longitudinal Recovery — `longitudinal.R`

Joins aggregated clinical features with a `sessions.csv`
(video → patient_id + session_date), computes session-to-session
deltas and percentage changes from baseline, and flags changes
exceeding 1 SD of the baseline session. Optionally overlays
external clinical scores on a secondary axis.

Outputs:

- `clinical_longitudinal_summary.csv` — values, deltas, flags
- `clinical_longitudinal_<patient_id>.png` — per-patient line plots

```{r longitudinal-summary}
long_csv <- file.path(out_dir, "clinical_longitudinal_summary.csv")
long_tbl <- read_csv_safe(long_csv)

if (is.null(long_tbl)) {
  cat("_No `clinical_longitudinal_summary.csv` found. Run:_ ",
      "`Rscript analysis/longitudinal.R output/ sessions.csv`\n")
} else {
  per_patient <- long_tbl |>
    group_by(patient_id) |>
    summarise(
      n_sessions = n_distinct(session_date),
      n_flagged  = sum(flagged %in% TRUE),
      largest_change_feature = feature[which.max(abs(pct_change))],
      largest_pct_change = pct_change[which.max(abs(pct_change))],
      .groups = "drop"
    )
  knitr::kable(per_patient, digits = 1,
               caption = "Per-patient longitudinal summary")
}
```

```{r longitudinal-plots, results = "asis"}
long_pngs <- list_files_safe(out_dir, "^clinical_longitudinal_.*\\.png$")
if (length(long_pngs) > 0) {
  for (p in long_pngs) {
    cat(sprintf("### `%s`\n\n![](%s)\n\n", basename(p), p))
  }
}
```

---

# Metadata Templates & Validation

## `make_templates.R`

Scans `*_clinical.csv` files for unique video names and writes two
placeholder CSVs:

- `clinical_scores_template.csv` — `video`, `GRASSP`, `UEMS`, `SCIM`
- `sessions_template.csv` — `video`, `patient_id`, `session_date`

Fill these in and save as `clinical_scores.csv` /
`sessions.csv` for use with `clinical_correlation.R` and
`longitudinal.R`.

```{r templates}
for (f in c("clinical_scores_template.csv", "sessions_template.csv")) {
  p <- file.path(out_dir, f)
  d <- read_csv_safe(p)
  if (!is.null(d)) {
    cat(sprintf("**`%s`** — %d rows, columns: %s\n\n",
                f, nrow(d), paste(names(d), collapse = ", ")))
  }
}
```

## `validate_metadata.R`

Validates a completed `clinical_scores.csv` or `sessions.csv`
against the clinical feature CSVs. Checks that required columns are
present, videos match and are unique, dates parse to ISO 8601, and
scores are numeric. Exits 0 on success, 1 on errors. It writes no
files — run it from the shell:

```bash
Rscript analysis/validate_metadata.R clinical_scores.csv output/
Rscript analysis/validate_metadata.R sessions.csv       output/
```

---

# Session Information

```{r session-info}
sessionInfo()
```
