---
title: "Pose-Estimation Pipeline — Analysis Summary"
author: "Generated from analysis/*.R"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    df_print: paged
    fig_width: 10
    fig_height: 6
params:
  output_dir: "../output"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo       = FALSE,
  warning    = FALSE,
  message    = FALSE,
  fig.align  = "center",
  out.width  = "100%"
)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(readr)
  library(stringr)
  library(purrr)
  library(ggplot2)
  library(tibble)
  library(jsonlite)
  library(knitr)
})

out_dir <- normalizePath(params$output_dir, mustWork = FALSE)
has_out <- dir.exists(out_dir)

show_png <- function(path, caption = NULL) {
  if (file.exists(path)) {
    knitr::include_graphics(path)
  } else {
    cat(sprintf("_Not available: `%s`_\n\n", basename(path)))
  }
}

read_csv_safe <- function(path) {
  if (file.exists(path)) {
    suppressMessages(read_csv(path, show_col_types = FALSE))
  } else {
    NULL
  }
}

list_files_safe <- function(dir, pattern) {
  if (dir.exists(dir)) {
    list.files(dir, pattern = pattern, full.names = TRUE)
  } else {
    character()
  }
}
```

# Overview

This report consolidates outputs from the R analysis scripts in
`analysis/`. These scripts consume CSVs from the pose-estimation pipeline
(`pose_estimation.main` / `pose_estimation.run`). The report follows the
typical script run order:

1. **Pipeline quality** — `summary.R`, `timeseries.R`, `keypoint_detail.R`
2. **Landmark feature exploration** — `features.R`
3. **Run comparison** — `compare.R`
4. **Clinical kinematic features** — `clinical_features.R`
5. **Clinical exploration** — `explore_clinical.R`
6. **Temporal clinical analysis** — `temporal_clinical.R`
7. **Between-video comparison** — `compare_clinical.R`
8. **Dimensionality reduction** — `clinical_dimreduce.R`
9. **External-score correlation** — `clinical_correlation.R`
10. **Longitudinal recovery** — `longitudinal.R`
11. **Metadata templates & validation** — `make_templates.R`, `validate_metadata.R`

The report reads all existing outputs under `r out_dir`. It silently
skips sections when their inputs do not exist yet.

```{r inventory}
count_landmark_csvs <- function(dir) {
  if (!dir.exists(dir)) return(NA_integer_)
  files <- list.files(dir, pattern = "\\.csv$", full.names = TRUE)
  excl <- "(_metrics|_diag|_smooth|_clinical|_kp_detail|_summary|_feature_rank)\\.csv$"
  length(files[!str_detect(basename(files), excl)])
}

count_pattern <- function(dir, pattern, extra_exclude = NULL) {
  if (!dir.exists(dir)) return(NA_integer_)
  files <- list.files(dir, pattern = pattern, full.names = TRUE)
  if (!is.null(extra_exclude)) {
    files <- files[!str_detect(basename(files), extra_exclude)]
  }
  length(files)
}

inventory <- tibble(
  category = c(
    "Landmark CSVs (pipeline export)",
    "Metrics CSVs (quality metrics)",
    "Per-keypoint detail CSVs",
    "Per-run summary JSONs",
    "Per-frame clinical CSVs",
    "Per-window clinical CSVs"
  ),
  n_files = c(
    count_landmark_csvs(out_dir),
    count_pattern(out_dir, "_metrics\\.csv$"),
    count_pattern(out_dir, "_kp_detail\\.csv$"),
    count_pattern(out_dir, "_summary\\.json$"),
    count_pattern(out_dir, "_clinical\\.csv$", extra_exclude = "_windows\\.csv$"),
    count_pattern(out_dir, "_clinical_windows\\.csv$")
  )
)

knitr::kable(
  inventory |> select(category, n_files),
  col.names = c("Output category", "# files"),
  caption = "Output inventory"
)
```

---

# Pipeline Quality

## Per-run baseline report — `summary.R`

`summary.R` ingests each `*_metrics.csv` and produces a JSON summary
(`*_summary.json`) plus a console report covering:

- Detection rates (body, synthetic hands, re-crop hands)
- Confidence (body visibility median/IQR, hand flag medians)
- Jitter (body/hand median and p95, px)
- Smoothing delta (raw − smoothed, px)
- Carry-forward rate and mean streak length
- Bone-length correction and joint-angle clamp rate
- Hand-to-arm matching distances
- Inference time (median and p95, ms)

```{r summary-json-table}
json_files <- list_files_safe(out_dir, "_summary\\.json$")

if (length(json_files) == 0) {
  cat("_No `*_summary.json` files found. Run:_ `Rscript analysis/summary.R output/`\n")
} else {
  flatten_json <- function(j, prefix = "") {
    out <- list()
    for (name in names(j)) {
      key <- if (prefix == "") name else paste0(prefix, ".", name)
      val <- j[[name]]
      if (is.list(val) && !is.null(names(val))) {
        out <- c(out, flatten_json(val, key))
      } else {
        out[[key]] <- val
      }
    }
    out
  }

  rows <- map_dfr(json_files, \(f) {
    j <- jsonlite::fromJSON(f)
    flat <- flatten_json(j)
    video <- str_remove(basename(f), "_summary\\.json$")
    tibble(video = video, !!!flat)
  })

  knitr::kable(
    rows |>
      select(video, n_frames,
             any_of(c(
               "detection.body_detection_rate",
               "confidence.body_vis_median",
               "jitter_px.body_median",
               "jitter_px.body_p95",
               "carry_forward.body_carry_pct",
               "inference_ms.median",
               "inference_ms.p95"
             ))),
    caption = "Per-video baseline quality (from `*_summary.json`)"
  )
}
```

## Temporal diagnostics — `timeseries.R`

For every `*_metrics.csv`, `timeseries.R` generates five diagnostic
plots:

- `_detection_timeline.png` — frame-by-frame body/hand status
  (detected / carried / missing)
- `_jitter.png` — jitter over time with spike markers (> 3 × median)
- `_confidence.png` — body visibility + hand flag traces
- `_smoothing.png` — raw-vs-smoothed delta
- `_constraints.png` — bone-length and joint-angle correction history

```{r timeseries-examples, results = "asis"}
timeseries_pngs <- list_files_safe(out_dir, "_detection_timeline\\.png$")

if (length(timeseries_pngs) == 0) {
  cat("_No timeseries plots found. Run:_ `Rscript analysis/timeseries.R output/`\n")
} else {
  # Show the first example per plot type
  example <- str_remove(basename(timeseries_pngs[1]), "_detection_timeline\\.png$")
  cat(sprintf("The report shows diagnostic plots for **%s** (%d videos total):\n\n",
              example, length(timeseries_pngs)))
  for (suffix in c("_detection_timeline", "_jitter", "_confidence",
                   "_smoothing", "_constraints")) {
    p <- file.path(out_dir, paste0(example, suffix, ".png"))
    if (file.exists(p)) {
      cat(sprintf("### `%s`\n\n", basename(p)))
      cat(sprintf("![](%s)\n\n", p))
    }
  }
}
```

## Per-keypoint analysis — `keypoint_detail.R`

`keypoint_detail.R` runs after the pipeline exports `*_kp_detail.csv`
files with the `--metrics-detail` flag. It produces:

- `_jitter_heatmap.png` — keypoint × time jitter heatmap
- `_worst_keypoints.png` — top-20 keypoints by median jitter
- `_trajectory.png` — smoothed x/y trajectories for shoulders + hand wrists

```{r keypoint-detail, results = "asis"}
kp_pngs <- list_files_safe(out_dir, "_jitter_heatmap\\.png$")

if (length(kp_pngs) == 0) {
  cat("_No per-keypoint plots found. Export with `--metrics-detail`, then run:_ ",
      "`Rscript analysis/keypoint_detail.R output/`\n")
} else {
  example <- str_remove(basename(kp_pngs[1]), "_jitter_heatmap\\.png$")
  cat(sprintf("The report shows per-keypoint plots for **%s** (%d videos total):\n\n",
              example, length(kp_pngs)))
  for (suffix in c("_worst_keypoints", "_jitter_heatmap", "_trajectory")) {
    p <- file.path(out_dir, paste0(example, suffix, ".png"))
    if (file.exists(p)) {
      cat(sprintf("### `%s`\n\n![](%s)\n\n", basename(p), p))
    }
  }
}
```

---

# Landmark Feature Exploration — `features.R`

`features.R` loads landmark CSVs from `hands-arms` or `body` tracking
mode. It drops fully NA rows and median-imputes sporadic NAs. It produces:

- `_variance.png` — per-feature variance (log scale), coloured by part
- `_correlation.png` — clustered Pearson correlation heatmap
- `_pca_scree.png` — cumulative variance explained
- `_pca_biplot.png` — PC1 vs PC2 with top-10 loading arrows,
  points coloured by timestamp
- `_umap.png` / `_umap_video.png` — 2D UMAP embeddings
- `_feature_rank.csv` — variance / max-abs-correlation / top-PC loading table

```{r features-rank-table}
rank_csvs <- list_files_safe(out_dir, "_feature_rank\\.csv$")

if (length(rank_csvs) == 0) {
  cat("_No `*_feature_rank.csv` found. Run:_ `Rscript analysis/features.R output/`\n")
} else {
  rank <- read_csv_safe(rank_csvs[1])
  if (!is.null(rank)) {
    cat(sprintf("The table shows the top 10 features by variance from `%s`:\n\n",
                basename(rank_csvs[1])))
    knitr::kable(
      rank |> arrange(desc(variance)) |> head(10) |>
        select(any_of(c("column", "part", "keypoint", "coord",
                        "variance", "max_abs_corr", "top_pc", "pc_loading"))),
      digits = 3
    )
  }
}
```

```{r features-plots, results = "asis"}
var_pngs <- list_files_safe(out_dir, "_variance\\.png$")
if (length(var_pngs) > 0) {
  example <- str_remove(basename(var_pngs[1]), "_variance\\.png$")
  cat(sprintf("The report shows plots for **%s**:\n\n", example))
  for (suffix in c("_variance", "_correlation", "_pca_scree",
                   "_pca_biplot", "_umap")) {
    p <- file.path(out_dir, paste0(example, suffix, ".png"))
    if (file.exists(p)) {
      cat(sprintf("![](%s)\n\n", p))
    }
  }
}
```

---

# Run Comparison — `compare.R`

`compare.R` compares two pipeline runs from `*_summary.json` or
`*_metrics.csv`. It writes `comparison.csv`. If you provide metrics CSVs,
it also writes `comparison_jitter.png` and `comparison_detection.png`.
These plots contain overlaid distributions and a bar chart.

```{r compare-table}
comp_csv <- file.path(out_dir, "comparison.csv")
comp <- read_csv_safe(comp_csv)
if (is.null(comp)) {
  cat("_No `comparison.csv` found. Run:_ ",
      "`Rscript analysis/compare.R run_a.json run_b.json`\n")
} else {
  knitr::kable(comp |> arrange(desc(abs(pct_change))) |> head(15),
               caption = "Top 15 metrics by |percentage change|",
               digits = 3)
}
```

```{r compare-plots, results = "asis"}
for (f in c("comparison_jitter.png", "comparison_detection.png")) {
  p <- file.path(out_dir, f)
  if (file.exists(p)) cat(sprintf("![](%s)\n\n", p))
}
```

---

# Clinical Kinematic Features — `clinical_features.R`

`clinical_features.R` derives per-frame features from the landmark
CSVs: joint angles, reach, grasp aperture, and wrist / fingertip
displacement. It also derives smoothness features for 1 s windows with
50 % overlap. These features include spectral arc length and wrist
velocity mean / peak. `clinical_features.R` writes two files per input:

- `<stem>_clinical.csv` — per-frame features
- `<stem>_clinical_windows.csv` — per-window features

```{r clinical-features-inventory}
frame_files <- list_files_safe(out_dir, "_clinical\\.csv$")
frame_files <- frame_files[!str_detect(basename(frame_files), "_windows\\.csv$")]
win_files   <- list_files_safe(out_dir, "_clinical_windows\\.csv$")

if (length(frame_files) == 0) {
  cat("_No `*_clinical.csv` found. Run:_ ",
      "`Rscript analysis/clinical_features.R output/`\n")
} else {
  inv <- tibble(
    video = str_remove(basename(frame_files), "_clinical\\.csv$"),
    frame_rows = map_int(frame_files, \(f) nrow(read_csv_safe(f)))
  )
  if (length(win_files) > 0) {
    inv <- inv |>
      left_join(
        tibble(
          video = str_remove(basename(win_files), "_clinical_windows\\.csv$"),
          window_rows = map_int(win_files, \(f) nrow(read_csv_safe(f)))
        ),
        by = "video"
      )
  } else {
    inv$window_rows <- NA_integer_
  }
  knitr::kable(inv, caption = "Clinical feature files (per video)")
}
```

---

# Clinical Exploration — `explore_clinical.R`

`explore_clinical.R` loads all per-frame and per-window clinical CSVs.
It writes four diagnostic plots and a console summary. The summary includes
row counts, per-column NA rates, and data-quality warnings. The warnings
cover features > 50 % NA, zero-variance columns, and empty videos.

```{r explore-plots, results = "asis"}
for (f in c("all_clinical_distributions.png",
            "all_clinical_na_heatmap.png",
            "all_clinical_boxplots.png",
            "all_clinical_window_distributions.png")) {
  p <- file.path(out_dir, f)
  if (file.exists(p)) {
    cat(sprintf("### `%s`\n\n![](%s)\n\n", f, p))
  }
}
```

---

# Temporal Clinical Analysis — `temporal_clinical.R`

For every video with ≥ 10 per-frame rows, `temporal_clinical.R`
produces a multi-panel time-series plot. The plot overlays the left and
right sides. When available, it includes a window-level SAL / velocity
panel. A cross-video overview plot compares normalised reach across all
qualifying videos.

```{r temporal-plots, results = "asis"}
ts_pngs <- list_files_safe(out_dir, "_clinical_timeseries\\.png$")
overview <- file.path(out_dir, "all_clinical_timeseries_overview.png")

if (file.exists(overview)) {
  cat("### Cross-video overview\n\n")
  cat(sprintf("![](%s)\n\n", overview))
}

if (length(ts_pngs) > 0) {
  cat(sprintf("### Example per-video time series (%d total)\n\n",
              length(ts_pngs)))
  cat(sprintf("![](%s)\n\n", ts_pngs[1]))
}

if (length(ts_pngs) == 0 && !file.exists(overview)) {
  cat("_No temporal plots found. Run:_ ",
      "`Rscript analysis/temporal_clinical.R output/`\n")
}
```

---

# Between-Video Comparison — `compare_clinical.R`

`compare_clinical.R` aggregates per-frame clinical features for each
video. It uses mean / median / SD / min / max. It aggregates window
features using mean / SD. It produces a wide summary table
(`all_clinical_video_summary.csv`) and two plots:

- `all_clinical_radar.png` — parallel-coordinate plot of
  z-scored feature means
- `all_clinical_heatmap.png` — clustered z-score heatmap

`compare_clinical.R` prints outlier videos (|z| > 2 on any feature
mean) to the console.

```{r compare-clinical-summary}
video_summary <- read_csv_safe(file.path(out_dir, "all_clinical_video_summary.csv"))

if (is.null(video_summary)) {
  cat("_No `all_clinical_video_summary.csv` found. Run:_ ",
      "`Rscript analysis/compare_clinical.R output/`\n")
} else {
  cat(sprintf("**The summary contains %d videos × %d aggregated feature columns.**\n\n",
              nrow(video_summary), ncol(video_summary) - 1))

  mean_cols <- names(video_summary)[str_ends(names(video_summary), "__mean")]
  if (length(mean_cols) > 0) {
    z_df <- video_summary |>
      select(video, all_of(mean_cols)) |>
      mutate(across(all_of(mean_cols), \(x) {
        s <- sd(x, na.rm = TRUE)
        if (is.na(s) || s == 0) return(rep(NA_real_, length(x)))
        (x - mean(x, na.rm = TRUE)) / s
      }))

    outliers <- z_df |>
      pivot_longer(-video, names_to = "feature", values_to = "z") |>
      filter(!is.na(z), abs(z) > 2) |>
      mutate(feature = str_remove(feature, "__mean$")) |>
      arrange(desc(abs(z)))

    if (nrow(outliers) > 0) {
      knitr::kable(outliers |> head(20),
                   digits = 2,
                   caption = "Outlier videos (|z| > 2 on any feature mean, top 20)")
    } else {
      cat("_No videos exceed |z| > 2 on any feature._\n\n")
    }
  }
}
```

```{r compare-clinical-plots, results = "asis"}
for (f in c("all_clinical_radar.png", "all_clinical_heatmap.png")) {
  p <- file.path(out_dir, f)
  if (file.exists(p)) {
    cat(sprintf("### `%s`\n\n![](%s)\n\n", f, p))
  }
}
```

---

# Dimensionality Reduction — `clinical_dimreduce.R`

`clinical_dimreduce.R` applies PCA and UMAP to the per-video aggregated
clinical features. It writes:

- `all_clinical_pca_scree.png` — variance explained per PC
- `all_clinical_pca_biplot.png` — PC1/PC2 scatter with top loading arrows
- `all_clinical_umap.png` — 2D UMAP
- `all_clinical_pca_loadings.csv` — feature loadings on first 5 PCs

```{r dimreduce-loadings}
loadings <- read_csv_safe(file.path(out_dir, "all_clinical_pca_loadings.csv"))

if (is.null(loadings)) {
  cat("_No PCA loadings found. Run:_ ",
      "`Rscript analysis/clinical_dimreduce.R output/`\n")
} else {
  knitr::kable(
    loadings |> head(10) |>
      select(any_of(c("feature_short", "PC1", "PC2", "PC3", "abs_PC1"))),
    digits = 3,
    caption = "Top 10 features loading on PC1"
  )
}
```

```{r dimreduce-plots, results = "asis"}
for (f in c("all_clinical_pca_scree.png",
            "all_clinical_pca_biplot.png",
            "all_clinical_umap.png")) {
  p <- file.path(out_dir, f)
  if (file.exists(p)) {
    cat(sprintf("### `%s`\n\n![](%s)\n\n", f, p))
  }
}
```

---

# External-Score Correlation — `clinical_correlation.R`

`clinical_correlation.R` joins aggregated clinical features with an
external `clinical_scores.csv` on `video`. The features use mean / median /
SD / min / max per video. It computes pairwise Pearson and Spearman
correlations with Benjamini–Hochberg FDR correction. It writes:

- `clinical_correlation_table.csv` — tidy correlation results
- `clinical_correlation_matrix.png` — Spearman heatmap with
  significance stars
- `clinical_scatter_top.png` — scatter plots for top 6 pairs

```{r correlation-table}
cor_tbl <- read_csv_safe(file.path(out_dir, "clinical_correlation_table.csv"))

if (is.null(cor_tbl)) {
  cat("_No `clinical_correlation_table.csv` found. Run:_ ",
      "`Rscript analysis/clinical_correlation.R output/ clinical_scores.csv`\n")
} else {
  top <- cor_tbl |>
    filter(!is.na(spearman_rho)) |>
    arrange(desc(abs(spearman_rho))) |>
    head(15)
  knitr::kable(top, digits = 3,
               caption = "Top 15 feature–score pairs by |Spearman ρ|")
}
```

```{r correlation-plots, results = "asis"}
for (f in c("clinical_correlation_matrix.png", "clinical_scatter_top.png")) {
  p <- file.path(out_dir, f)
  if (file.exists(p)) {
    cat(sprintf("### `%s`\n\n![](%s)\n\n", f, p))
  }
}
```

---

# Longitudinal Recovery — `longitudinal.R`

`longitudinal.R` joins aggregated clinical features with `sessions.csv`.
`sessions.csv` maps video → patient_id + session_date. The script computes
session-to-session deltas and percentage changes from baseline. It flags
changes that exceed 1 SD of the baseline session. It can overlay external
clinical scores on a secondary axis.

It writes:

- `clinical_longitudinal_summary.csv` — values, deltas, flags
- `clinical_longitudinal_<patient_id>.png` — per-patient line plots

```{r longitudinal-summary}
long_csv <- file.path(out_dir, "clinical_longitudinal_summary.csv")
long_tbl <- read_csv_safe(long_csv)

if (is.null(long_tbl)) {
  cat("_No `clinical_longitudinal_summary.csv` found. Run:_ ",
      "`Rscript analysis/longitudinal.R output/ sessions.csv`\n")
} else {
  per_patient <- long_tbl |>
    group_by(patient_id) |>
    summarise(
      n_sessions = n_distinct(session_date),
      n_flagged  = sum(flagged %in% TRUE),
      largest_change_feature = feature[which.max(abs(pct_change))],
      largest_pct_change = pct_change[which.max(abs(pct_change))],
      .groups = "drop"
    )
  knitr::kable(per_patient, digits = 1,
               caption = "Per-patient longitudinal summary")
}
```

```{r longitudinal-plots, results = "asis"}
long_pngs <- list_files_safe(out_dir, "^clinical_longitudinal_.*\\.png$")
if (length(long_pngs) > 0) {
  for (p in long_pngs) {
    cat(sprintf("### `%s`\n\n![](%s)\n\n", basename(p), p))
  }
}
```

---

# Metadata Templates & Validation

## `make_templates.R`

Scans `*_clinical.csv` files for unique video names and writes two
placeholder CSVs:

- `clinical_scores_template.csv` — `video`, `GRASSP`, `UEMS`, `SCIM`
- `sessions_template.csv` — `video`, `patient_id`, `session_date`

Fill these templates. Save them as `clinical_scores.csv` /
`sessions.csv`. Use them with `clinical_correlation.R` and `longitudinal.R`.

```{r templates}
for (f in c("clinical_scores_template.csv", "sessions_template.csv")) {
  p <- file.path(out_dir, f)
  d <- read_csv_safe(p)
  if (!is.null(d)) {
    cat(sprintf("**`%s`** — %d rows, columns: %s\n\n",
                f, nrow(d), paste(names(d), collapse = ", ")))
  }
}
```

## `validate_metadata.R`

`validate_metadata.R` validates a completed `clinical_scores.csv` or
`sessions.csv` against the clinical feature CSVs. It checks that files
contain the required columns. It checks that videos match and are unique.
It checks that dates parse to ISO 8601 and that scores are numeric.
`validate_metadata.R` exits 0 on success and 1 on errors. It writes no
files. Run it from the shell:

```bash
Rscript analysis/validate_metadata.R clinical_scores.csv output/
Rscript analysis/validate_metadata.R sessions.csv       output/
```

---

# Session Information

```{r session-info}
sessionInfo()
```
